Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-05-18 WSSSPE4_report/WSSSPE4_paper.tex
another typo
minor date formatting change

\begin{abstract}

This article summarizes motivations, organization, and activities of the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4). The WSSSPE series promotes sustainable research software by positively impacting principles and best practices, careers, learning, and credit. This article includes the mission and vision statements that were drafted at the workshop and finalized shortly after it; the keynote and idea papers, position papers, experience papers, demos, and lightning talks presented during the workshop; and a panel discussion on best practices. The main part of the article discusses the set of working groups that formed during the meeting, along with contact information for readers who may want to join a group. Finally, it discusses a survey of the workshop attendees.

@@

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held over 2 1/2 days on 12--14 September 2016 in Manchester, England, UK.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous events in the WSSSPE series include
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
held in conjunction with the SC13 conference;
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}}, a
focused workshop organized jointly with the SciPy
conference\footnote{\url{https://conference.scipy.org/scipy2014/participate/wssspe/}};
WSSSPE2\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}}~\cite{WSSSPE2-pre-report,WSSSPE2},
held in conjunction with SC14;
WSSSPE2.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}}, a
focused workshop organized again jointly with
SciPy\footnote{\url{http://scipy2015.scipy.org/ehome/115969/286469/}};
and WSSSPE3\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe3/}}~\cite{WSSSPE3},
held in Boulder, Colorado, USA.
The first WSSSPE workshop was named
``Working towards
Sustainable Software for Science: Practice and Experiences,'' which remains the meaning
of the WSSSPE group, but the workshops after that were named
``Workshop on Sustainable
Software for Science: Practice and Experiences.'' Together these reflect
that WSSSPE is both a community and a set of workshops.

@@

The WSSSPE4 workshop included multiple mechanisms for participation and
encouraged team building around solutions. WSSSPE4 strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional people to attend and participate. A subset of the
organizing committee reviewed 44 applications for travel support and
competitively selected the awardees, including 11 students and five early-career
researchers. In addition, the travel award subcommittee tried to increase the
diversity of applicants by forwarding the notice of travel support to
organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
Django Girls\footnote{\url{https://djangogirls.org}},
Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}
%
WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

At the workshop, we introduced a Code of Conduct (CoC).%
\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/code-of-conduct/}}
The CoC was conceived for the workshop and a guideline for the community of scientists which WSSSPE
supports, including their personal and online interactions (e.g., on
Twitter, in email lists, in the Slack team). The WSSSPE4 CoC is based on the
FORCE11 conference CoC~\cite{FORCE11:CoC}, which is in turn based on the Code4Lib
CoC~\cite{Code4Lib:CoC}.
The main guidelines of the CoC are:
\begin{quote}
    WSSSPE events are community events intended for networking and collaboration
    as well as learning. We value the participation of every member of the
    community and want all attendees to have an enjoyable and fulfilling
    experience. Accordingly, all attendees are expected to show respect and
    courtesy to other attendees throughout the event and in interactions online
    associated with the event.

@@

WSSSPE4 was based on the work done at previous WSSSPE meetings, but aimed
at producing working groups that better continued working after the workshop ended.
In addition, based on feedback after WSSSPE3, it became clear that attendees
had two distinct motivations:  to make a better future
for research software, and to immediately do better research software development.
This led to the idea of WSSSPE4 being partially divided into two tracks:

\begin{quote}
    \textbf{Track 1 -- Building a sustainable future for open-use research
    software}: defining a vision of the future of open-use
    research software, and in the workshop, initiating the activities that are
    needed to get there.

    \noindent \textbf{Track 2 -- Practices \& experiences in sustainable scientific software}:
    improving the quality of today's research software and the
    experiences of its developers by sharing best practices and experiences.
\end{quote}

The call for participation requested:
idea papers, implementable proposals related to Track~1;
position papers, longer, not previously published papers
discussing what we can do to improve sustainable scientific
software in the short term;
experience papers, longer papers that discuss current
experiences and how they have been used to improve the quality of
today's research software and/or the experiences of its developers;
demos, brief papers describing a tool or
process that would be demonstrated that improves the quality of today's research
software and\slash or the experiences of its developers; and
extended abstracts describing lightning talks,
where each author could make a brief statement about work that either had been
done or was needed.

Suggested contribution topics included:

\begin{itemize}
\renewcommand{\labelenumi}{\textbf{\theenumi}.}
\setlength{\rightmargin}{1em}

\item Development and Community:
best practices for developing sustainable software;
models for funding specialist expertise in software collaborations;
software tools that aid sustainability;
academia/industry interaction;
refactoring/improving legacy scientific software;
engineering design for sustainable software;
metrics for the success of scientific software; and
adaptation of mainstream software practices for scientific software

\item Professionalization:
career paths;
RSE as a brand;
RSE outside of the UK or Europe; and
Increase incentives in publishing, funding and promotion for better software

\item Training:
taining for developing sustainable software; and 
curriculum for software sustainability

\item Credit:
making the existing credit and citation ecosystem work better for software;
future credit and citation ecosystem;
software contributions as a part of tenure review;
case studies of receiving credit for software contributions; and
awards and recognition that encourage sustainable software

\item Software publishing:
journals and alternative venues for publishing software; and
review processes for published software

\item Software discoverability/reuse:
proposals and case studies

\item Reproducibility and testing:
reproducibility in conferences and journals; and
best practices for code testing and code review

@@

Submissions to WSSSPE4 comprised
19 lightning talks,
4 idea papers,
3 position paper,
5 experience papers,
and
3 demos.
Most of the submissions were accepted, since the goal of WSSSPE is always to
take in as many relevant inputs as possible, and to encourage their authors to
participate in sharing and implementing their ideas.
Specifically,
19 lightning talks,
4 idea papers,
3 position papers,
2 experience papers,
and
3 demos
were accepted, as mentioned in Section~\ref{sec:presentations}. (Two of the submitted lightning talks and one submitted experience paper were rejected, while two more submitted experience papers
were accepted as lightning talks.)
The papers and lightning talks have been published as a volume in the CEUR Workshop Proceedings~\cite{WSSSPE4-proceedings}.

@@

Going into WSSSPE4, WSSSPE had no formal mission or vision statement.
The organizers developed a strawhorse, which was presented to the participants early in the meeting.
The presentation included guidelines as well as examples from other similar communities.
The guidelines were that in general, an organization's mission should stand the test of time and state what the organization does; its vision should imagine what the world would look like if the organization is successful; and from which, focus areas could be used to establish the scope of the organization along with its goals.

@@

A final draft was put on GitHub and advertised
via the WSSSPE mailing list, Facebook group, and Twitter.
After two weeks without suggested changes, the final statements, as shown below, were added
to the WSSSPE web page (\url{http://wssspe.researchcomputing.org.uk/about-wssspe/}).

@@

\textbf{Definitions:}
\emph{Sustainable software} has the capacity to endure such that it will continue to
be available in the future, on new platforms, meeting new needs.
The \emph{research software lifecycle} includes:
acquiring and assembling resources (including funding and people) into teams and communities,
developing software,
using software,
recognizing contributions to and of software,
and
maintaining software.

@@

After most of the lightning talks and other presentations, WSSSPE used three areas of a large room to let
attendees use sticky flip charts and sticky notes on the walls to suggest a vision on any aspect of the work, a gap or challenge, or a project idea.
Next, attendees organically formed groups around the flip charts, and 12 groups emerged.
Summaries of each group's activities follow; for more details
about the group discussions see \cite{WSSSPE4-report}.

@@

%from Santiago
There is a productivity bottleneck---yet to be solved---in HPC from the human
perspective, first identified in the 1980s~\cite{barstow1982automatic}. Of all
domain-specific scientists, only a percentage of those use simulations due to
perceptions mostly related to the difficulty of using and developing those tools.
XSEDE and other resources have become a gateway for successfully increasing the
basis of HPC scientific users by facilitating their access to infrastructure and
tools~\cite{towns2014xsede}, but the development challenge remains.
Of all scientists who use simulations, only a small percentage know how
to develop prototypes or full applications that can be later scaled by computer
scientists and engineers. Writing new parallel applications is unusual for
domain scientists and thus has created a complex dependency on specialized
scientific programmers, who are scarce~\cite{post2005computational}. However,
the capacity to quickly envision and prototype applications by domain
experts---later to be fully implemented into by scientific programmers---is critical for
acquiring maturity and proficiency in the grand challenges that Exascale
attempts to solve, and moreover for allowing scientists to develop their own
ideas quickly and productively~\cite{vinter2015prototyping}. The latter, in
addition, would increase service utilization time in HPC systems, a critical
measure of efficiency. Considering the existing code base in scientific
computing, software infrastructures that allow a clean transition from legacy
systems to new Exascale platforms while preserving flexible prototyping
towards the future are central~\cite{hwu2015transitioning}.

@@

The primary objective of this working group, which included people who had previously been working on the CodeMeta project, is to find ways to help the project come to fruition.
%
The working group plans to engage with those already working on CodeMeta, to examine the existing CodeMeta crosswalk table to see what improvements and additions might be made, and to determine how to engage the community and provide ongoing social engagement and structure. Further, the group wants to assist in the implementation of the specifications and, by providing an outsider's view, contribute suggestions for more understandable project documentation. Finally, the group seeks a better way or ways to present the crosswalk table to make it more easily understood and consumable by research software communities.
%
While at WSSSPE4, the group greatly expanded the CodeMeta project README file to include a description of the project geared to those with little or no prior knowledge of the project, a list of contributors, information on how one can get involved, a brief project history and who is managing the project, and links to additional information.

@@

There is more and more academic research being done on topics related to software sustainability, including work on software engineering practices and management of open source projects. However, academic research in general is often siloed for many reasons, and work on topics relevant to software sustainability is no exception. There are many academic studies and projects which may be relevant for practitioners working in this area, and many projects and initiatives by practitioners that may be relevant for social scientists. However, there is a gap between these two in research and practice.
%
This working group is motivated by the goal of building better connections between academic researchers who are studying topics in or relating to software sustainability with practitioners, managers, and administrators who are working in the area of software sustainability.
%
 In any domain, bringing research and practice closer together a mutually beneficial goal, but also has its challenges.
The group aims to give social scientists and practitioners working on scientific software, software sustainability, and open source and a better understanding of each other's work, as well as help them connect and coordinate on specific projects of mutual interest.
%
Furthermore, the group also recognize that academic social science and research software engineering are not monolithic, and there is a need to connect people who care about research-driven best practices inside of these two domains with each other as well.

@@

This working group was motivated by the perceived prevalence of
so-called ``hidden code'' in scientific communities: code written by
individual researchers in an unsustainable way that is never shared
with the larger community. Participants recalled their own experience
working with colleagues who write hidden code or inheriting
unsustainable hidden code from collaborators.
%
The question is, then, how to prevent researchers from writing hidden
code? The participants hypothesized that the best strategy is to catch
researchers while they are still in training and teach them software
best-practices. Therefore, this working group's goal
is to develop courses on software best practices aimed at
undergraduate students studying domain science. The program might be
similar to a Software Carpentry or Data Carpentry workshop but
focused for domain scientists.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Coordinating access to continuous integration (CI) for research softwaree}
\label{sec:access}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This group believes that the authors of scientific software provide important
services to departments that are no less than strictly disciplinary
research. Consequently, leveling the playing field with more
disciplinary candidates for hires, tenure, and promotion requires that
letter writers be aware of how their letters will be read by
committees. It also requires that committees be aware that such
letters often look different and may provide a different perspective
of how a candidate's achievements should be assessed. For example, in
mathematics a typical candidate would be evaluated on the difficulty
and depth of the statements she may have proven in their papers. On
the other hand, an author of mathematical software would likely be
evaluated based on the impact of her software, or the number of citations of
the publications that describe it. She may also be evaluated by how
widely the software is used \textit{outside} mathematics, a criterion
that is rarely used for more disciplinary mathematicians.

@@

All participants present on the last day of the meeting, September 14, 2016, were asked to complete the online survey administered through Qualtrics\footnote{\url{https://www.qualtrics.com}.}.
44 responses were recorded, with one respondent indicating that they took the survey twice.
The survey URL was also later distributed to attendees by email and advertised on twitter for attendees who were not there on the last day to have a chance to respond, and one additional response was recorded after the meeting.
Hence, the total number of unique individuals responding to the survey is likely 44.
The survey contained nine multiple choice and open-text answer questions.
Text responses have been alphabetize to preserve anonymity.
The survey had a 100\% completion rate.

>>>>

\begin{abstract}
This report records and discusses the Fourth Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE4). The report
includes a description of the keynote presentation of the workshop, the
mission and vision statements that were drafted at the workshop and
finalized shortly after it, a set of idea papers, position papers, experience
papers, demos, and lightning talks, and a panel discussion.  The main part of the report
covers the set of working groups that formed during the meeting, along with contact information for readers who may want to join
a group. Finally, we present results from a survey of the workshop attendees.

@@

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held over 2 1/2 days on 12--14 September 2016 in Manchester, England, UK.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous events in the WSSSPE series include
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
held in conjunction with the SC13 conference;
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}}, a
focused workshop organized jointly with the SciPy
conference\footnote{\url{https://conference.scipy.org/scipy2014/participate/wssspe/}};
WSSSPE2\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}}~\cite{WSSSPE2-pre-report,WSSSPE2},
held in conjunction with SC14;
WSSSPE2.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}}, a
focused workshop organized again jointly with
SciPy\footnote{\url{http://scipy2015.scipy.org/ehome/115969/286469/}};
and WSSSPE3\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe3/}}~\cite{WSSSPE3},
held in Boulder, Colorado, USA.
Note that the first WSSSPE workshop was named
``Working towards
Sustainable Software for Science: Practice and Experiences,'' which remains the meaning
of the WSSSPE group, but the workshops after that were named
``Workshop on Sustainable
Software for Science: Practice and Experiences.'' Together these reflect
that WSSSPE is both a community and a set of workshops.

@@

The WSSSPE4 workshop included multiple mechanisms for participation and
encouraged team building around solutions. WSSSPE4 strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional people to attend and participate. A subset of the
organizing committee reviewed 44 applications for travel support and
competitively selected the awardees, including 11 students and five early-career
researchers. In addition, the travel award subcommittee tried to increase the
diversity of applicants by forwarding the notice of travel support to
organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
Django Girls\footnote{\url{https://djangogirls.org}},
Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}

WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

At the workshop, we also introduced a Code of Conduct (CoC).%
\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/code-of-conduct/}}
The CoC was conceived for the workshop itself; however, we intend it
to also be a guideline for the community of scientists which WSSSPE
supports, and their personal and online interactions (e.g., on
Twitter, in email lists, in the Slack team). The WSSSPE4 CoC is based on the
FORCE11 conference CoC~\cite{FORCE11:CoC}, in turn based on the Code4Lib
CoC~\cite{Code4Lib:CoC}.
The main guidelines of the CoC are:
\begin{quote}
    WSSSPE events are community events intended for networking and collaboration
    as well as learning. We value the participation of every member of the
    community and want all attendees to have an enjoyable and fulfilling
    experience. Accordingly, all attendees are expected to show respect and
    courtesy to other attendees throughout the event and in interactions online
    associated with the event.

@@

WSSSPE4 was based on the work done at WSSSPE1, WSSSPE2, and WSSSPE3, but aimed
at producing working groups that better continued working after the workshop ended.
In addition, based on feedback after WSSSPE3, it became clear that WSSSPE attendees
had two different motivations in participating.  One motivation was to make a better future
for research software, and the other was to immediately do better research software development.
This led to the idea of WSSSPE4 being partially divided into two tracks:

\begin{quote}
    \textbf{Track 1 -- Building a sustainable future for open-use research
    software} has the goals of defining a vision of the future of open-use
    research software, and in the workshop, initiating the activities that are
    needed to get there. The idea of this track is to first think about where
    we want to be 5 to 10 years from now, without being too concerned with
    where we are today, and then to determine how we can move to this future.

    \noindent \textbf{Track 2 -- Practices \& experiences in sustainable scientific software}
    has the goal of improving the quality of today's research software and the
    experiences of its developers by sharing best practices and experiences.
    This track is focused on the current state of scientific software and what
    we can do to improve it in the short term, starting with where we are today.
\end{quote}

The call for participation requested
idea papers, which present implementable proposals related to Track~1;
position papers, which are longer, not previously published papers related to
Track~2 specifically discussing what we can do to improve sustainable scientific
software in the short term, starting with where we are today;
experience papers, longer papers related to Track~2 that discuss current
practices and experiences and how they have been used to improve the quality of
today's research software and/or the experiences of its developers;
demos, brief papers describing a tool or
process that would be demonstrated, relevant to Track~2 that improves the quality of today's research
software and\slash or the experiences of its developers; and
extended abstracts describing lightning talks,
where each author could make a brief statement about work that either had been
done or was needed.

The following list of contribution topics was suggested, though this was not intended to limit potential submissions:

%
%\begin{quote}
\begin{itemize}
\renewcommand{\labelenumi}{\textbf{\theenumi}.}
\setlength{\rightmargin}{1em}

\item Development and Community
\begin{itemize}
    \item Best practices for developing sustainable software
    \item Models for funding specialist expertise in software collaborations
    \item Software tools that aid sustainability
    \item Academia/industry interaction
    \item Refactoring/improving legacy scientific software
    \item Engineering design for sustainable software
    \item Metrics for the success of scientific software
    \item Adaptation of mainstream software practices for scientific software
\end{itemize}

\item Professionalization
\begin{itemize}
    \item Career paths
    \item RSE as a brand
    \item RSE outside of the UK or Europe
    \item Increase incentives in publishing, funding and promotion for better software
\end{itemize}

\item Training
\begin{itemize}
    \item Training for developing sustainable software
    \item Curriculum for software sustainability
\end{itemize}

\item Credit
\begin{itemize}
    \item Making the existing credit and citation ecosystem work better for software
    \item Future credit and citation ecosystem
    \item Software contributions as a part of tenure review
    \item Case studies of receiving credit for software contributions
    \item Awards and recognition that encourage sustainable software
\end{itemize}

\item Software publishing
\begin{itemize}
    \item Journals and alternative venues for publishing software
    \item Review processes for published software
\end{itemize}

\item Software discoverability/reuse
\begin{itemize}
    \item Proposals and case studies
\end{itemize}

\item Reproducibility and testing
\begin{itemize}
    \item Reproducibility in conferences and journals
    \item Best practices for code testing and code review
\end{itemize}

@@

Submissions to WSSSPE4 comprised
19 lightning talks,
4 idea papers,
3 position paper,
5 experience papers,
and
3 demos.
Most of the submissions were accepted, since the goal of WSSSPE is always to
take in as many relevant inputs as possible, and to encourage their authors to
participate in sharing and implementing their ideas.
Specifically,
19 lightning talks,
4 idea papers,
3 position papers,
2 experience papers,
and
3 demos
were accepted, as mentioned in Section~\ref{sec:presentations}. (Two of the submitted lightning talks and one submitted experience paper were rejected, while two more submitted experience papers
were accepted as lightning talks.)
The paper and lightning talks have been published as a volume in the CEUR Workshop Proceedings~\cite{WSSSPE4-proceedings}.

@@

Going into WSSSPE4, the WSSSPE organization had not had a formal mission or vision statement.
The organizers developed a strawhorse, which was presented to the participants early in the meeting.
The presentation included guidelines as well as examples from other similar communities.
The guidelines were that in general, an organization's mission should stand the test of time and state what is wrong and how the organization is going to fix it; its vision should imagine what the world look like if the organization is successful; and from which, focus areas could be use to establish the scope of the organization along with its goals.

@@

A final draft was put on GitHub (\url{https://github.com/WSSSPE/mission-vision}) and advertised
via the WSSSPE mailing list, Facebook group, and Twitter.
After two weeks without suggested changes, the final statements, as shown below, were added
to the WSSSPE web page (\url{http://wssspe.researchcomputing.org.uk/about-wssspe/}).

@@

\textbf{Definitions:}
\emph{Sustainable software} has the capacity to endure such that it will continue to
be available in the future, on new platforms, meeting new needs.
The \emph{research software lifecycle} includes:
\begin{itemize}
\item acquiring and assembling resources (including funding and people) into teams and communities
\item developing software
\item using software
\item recognizing contributions to and of software
\item maintaining software
\end{itemize}

@@

After most of the lightning talks and other presentations, WSSSPE used three areas of a large room to let
attendees use sticky flip charts and sticky notes on the walls to suggest: (1) a vision on any aspect of the work, (2) a gap or challenge, or (3) a project idea.
Next, attendees organically formed groups around the flip charts on the walls.  This led to the 12 groups that are discussed in the rest of this section.
Summaries of each group's activities follow; for more details
about the group discussions see \cite{WSSSPE4-report}.

@@

%from Santiago
There is a productivity bottleneck---yet to be solved---in HPC from the human
perspective, first identified in the 1980s~\cite{barstow1982automatic}. Of all
domain-specific scientists, only a percentage of those use simulations due to
perceptions mostly related to the difficulty of using and developing those tools.
XSEDE and other resources have become a gateway for successfully increasing the
basis of HPC scientific users by facilitating their access to infrastructure and
tools~\cite{towns2014xsede}, but the development challenge remains.
Of all scientists who use simulations, only a small percentage know how
to develop prototypes or full applications that can be latter scaled by computer
scientists and engineers. Writing new parallel applications is unusual for
domain scientists and thus has created a complex dependency on specialized
scientific programmers, who are scarce~\cite{post2005computational}. However,
the capacity to quickly envision and prototype applications by domain
experts---later to be fully implemented into by scientific programmers---is critical for
acquiring maturity and proficiency in the grand challenges that Exascale
attempts to solve, and moreover for allowing scientists to develop their own
ideas quickly and productively~\cite{vinter2015prototyping}. The latter, in
addition, would increase service utilization time in HPC systems, a critical
measure of efficiency. Considering the existing code base in scientific
computing, software infrastructures that allow a clean transition from legacy
systems to new Exascale platforms while preserving flexible prototyping
towards the future are central~\cite{hwu2015transitioning}.

@@

The primary objective of this working group, which included people who had previously been working on the CodeMeta project, is to find ways to help the project come to fruition.
%
The working group plans to engage with those already working on CodeMeta, to examine the existing CodeMeta crosswalk table to see what improvements and additions might be made, and to determine how to engage the community and provide ongoing social engagement and structure. Further, the group wants to assist in the implementation of the specifications and, by providing an outsider's view, contribute suggestions for more understandable project documentation. Finally, the group seeks a better way or ways to present the crosswalk table to make it more easily understood and consumable by research software communities.
%
While at WSSSPE4, the group greatly expanded the CodeMeta project README file\footnote{\url{https://github.com/codemeta/codemeta/blob/master/README.md}} to include a description of the project geared to those with little or no prior knowledge of the project, a list of contributors, information on how one can get involved, a brief project history and who is managing the project, and links to additional information.

@@

There is more and more academic research being done on topics related to software sustainability, including work on software engineering practices and management of open source projects. However, academic research in general is often siloed for many reasons, and work on topics relevant to software sustainability is no exception. There are many academic studies and projects which may be relevant for practitioners working in this area, and many projects and initiatives by practitioners that may be relevant for social scientists. However, there is a gap between these two research and practice.
%
This working group is motivated by the goal of building better connections between academic researchers who are studying topics in or relating to software sustainability with practitioners, managers, and administrators who are working in the area of software sustainability.
%
 In any domain, bringing research and practice closer together a mutually beneficial goal, but also has its challenges.
The group aims to give social scientists and practitioners working on scientific software, software sustainability, and open source and a better understanding of each other's work, as well as help them connect and coordinate on specific projects of mutual interest.
%
Furthermore, the group also recognize that academic social science and research software engineering are not monolithic, and there is a need to connect people who care about research-driven best practices inside of these two domains with each other as well.

@@

This working group was motivated by the perceived prevalence of
so-called ``dark code'' in scientific communities: code written by
individual researchers in an unsustainable way that is never shared
with the larger community. Participants recalled their own experience
working with colleagues who write dark code or inheriting
unsustainable dark code from collaborators.
%
The question is, then, how to prevent researchers from writing dark
code? The participants hypothesized that the best strategy is to catch
researchers while they are still in training and teach them software
best-practices. Therefore, this working group's goal
is to develop courses on software best practices aimed at
undergraduate students studying domain science. The program might be
similar to a Software Carpentry or Data Carpentry workshop but
focused for domain scientists.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Coordinating access to CI for research software}
\label{sec:access}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This group believes that the authors of scientific software provide important
services to departments that are no less than strictly disciplinary
research. Consequently, in order to level the playing field with more
disciplinary candidates for hires, tenure, and promotion requires that
letter writers be aware of how their letters will be read by
committees. It also requires that committees be aware that such
letters often look different and may provide a different perspective
of how a candidate's achievements should be assessed. For example, in
mathematics a typical candidate would be evaluated on the difficulty
and depth of the statements she may have proven in their papers. On
the other hand, an author of mathematical software would likely be
evaluated based on the impact of her software, or the number of citations of
the publications that describe it. She may also be evaluated by how
widely the software is used \textit{outside} mathematics, a criterion
that is rarely used for more disciplinary mathematicians.

@@

All participants present on the last day of the meeting, 14 September 2016, were asked to complete the online survey administered through Qualtrics\footnote{\url{https://www.qualtrics.com}.}.
44 responses were recorded, with one respondent indicating that they took the survey twice.
The survey URL was also later distributed to attendees by email and advertised on twitter for attendees who were not there on the last day to have a chance to respond, and one additional response was recorded after the meeting.
Hence, the total number of unique individuals responding to the survey is likely 44.
The survey contained nine multiple choice and open-text answer questions.
Text responses have been alphabetize to preserve anonymity.
The survey had a 100\% completion rate.
